% $POSTERS_SESSION_PVI_p6
P-VI-3

\atitle{The Ultra-Fast Flash Observatory Program : Pathfinder and Beyond}

\bigskip

\authors{B. Grossan [1,2], I. H. Park [3,] , A. J. Castro-Tirado [4], P. Chen [5], P. Connell [6], S. Dagoret-Campagne [7], C. Eyles [6],  A. Iyudin [1], J. Jeon  [3], S. Jeong [3], A. Jung [3], J. Karczmarczyk [8], J. Kim [3], S. H. Kim [9],  J. Lee [3],  H. Lim [10], E. Linder [4], T. Liu [5], Alfred Ming-huey [5],   G. Na [3], J. W. Nam [5], Vladilav Osdedlo [1], M. Panasyuk [1], V. Reglero [6], J. Ripa [10], G. Smoot [1,10,11], J Szabelski [8], J. E. Suh [3], U. Seljak [10],  S. Svertilov [1], C. De La Taille [7], N. Vedenkin [12], J. Yang [10], I. Yashin [12]}

\affiliation{[1] Extreme Universe Lab., Moscow State University , [2] UC Berkeley Space Sciences Lab., [3] Research Center for MEMS Space Telescope, Ewha Womans Univ. , [4] Instituto de Astrofisica de Andalucia, [5] Leung Center for Cosmology and Particle Astrophysics, National Taiwan University, [6] Univ. of Valencia, [7] Laboratoire de l'Accelerateur Lineaire, Univ. de Paris Sud - 11 , [8] Andrzej Soltan Institute for Nuclear Science Cosmic Ray Lab. , [9] Yonsei Univ., [10] Institute for the Early Universe,  Ewha Womans Univ., [11] Univ. California, Berkeley}

\bigskip

\noindent This talk describes the current status of the Ultra-Fast Flash Observatory (UFFO) program, a series of gamma-ray burst (GRB) space missions to understand the rise phase (prompt) optical/IR emission of GRB.  We report the status of the UFFO Pathfinder, a small prototype system with an X-ray coded mask camera (191 cm$^{2}$ collecting area) coupled to an optical telescope (10 cm aperture) for fast (< 1 s pointing) X-ray trigger follow-up. The UFFO Pathfinder is scheduled to launch late this year and issue alerts to the community.  We also plan a full-size observatory for launch in 2015 with both optical and IR cameras, and possible additional instruments.  We describe the science drivers for our program, and discuss how fast optical response makes rich, new science possible.

\index{\tiny{Grossan, Bruce: \textit{The Ultra-Fast Flash Observatory Program : Pathfinder and Beyond}}}
